Faithful Variable Screening for High-Dimensional Convex Regression

### 2014-07-31 tex/scam/paper.tex
results update

% ENVIRONMENTS
\numberwithin{equation}{section}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheoremstyle{remark}{\topsep}{\topsep}%
     {\normalfont}% Body font
     {}           % Indent amount (empty = no indent, \parindent = para indent)
     {\bfseries}  % Thm head font
     {.}          % Punctuation after thm head
     {.5em}       % Space after thm head (\newline = linebreak)
     {\thmname{#1}\thmnumber{ #2}\thmnote{ #3}}% Thm head spec
\theoremstyle{remark}
\newtheorem{remark}{Remark}[section]
\newtheorem{example}{Example}[section]
\newtheorem{assumption}{Assumption}[section]
\newtheorem{definition}{Definition}[section]

@@

% John's macros
\def\X{\mathcal{X}}
\def\comma{\unskip,~}
\def\truep{p^*}
\def\div{\|\,}
\long\def\comment#1{}
\def\reals{{\mathbb R}}
\def\P{{\mathbb P}}
\def\E{{\mathbb E}}
\def\Cov{\mathop{\text{Cov}}}
\def\supp{\mathop{\text{supp}\kern.2ex}}
\def\argmin{\mathop{\text{\rm arg\,min}}}
\def\arginf{\mathop{\text{\rm arg\,inf}}}
\def\argmax{\mathop{\text{\rm arg\,max}}}
\let\tilde\widetilde
\def\csd{${}^*$}
\def\mld{${}^\dag$}
\def\dos{${}^\ddag$}
\def\W{\widetilde Y}
\def\Z{\widetilde X}
\let\hat\widehat
\let\tilde\widetilde
\def\given{{\,|\,}}
\def\ds{\displaystyle}
\def\bs{\backslash}
\def\1{{(1)}}
\def\2{{(2)}}
\def\pn{{(n)}}
\def\ip{{(i)}}
\def\Xbar{\overline{X}}
\def\except{\backslash}
\def\npn{\mathop{\textit{NPN\,}}}
\def\i{{(i)}}
\def\cE{{\mathcal{C}}}
\def\cM{{\mathcal{M}}}
\def\cF{{\mathcal{F}}}
\def\cP{{\mathcal{P}}}
\def\cG{{\mathcal{G}}}
\def\tr{\mathop{\text{tr}}}
\long\def\comment#1{}
\def\ti#1{#1}
\def\titi#1{\textit{#1}}
\def\cram{{\sc cram}}
\def\spam{{\small\sc SpAM}}
\def\diag{\mathop{\rm diag}}
\def\ones{\mathbf{1}}
\def\threebars{\mbox{$|\kern-.25ex|\kern-.25ex|$}}
\def\fatnorm#1{\threebars #1 \threebars}
\def\rank{\mathop{\rm rank}}
\def\S{\mathcal{S}}
\def\H{\mathcal{H}}
\def\K{{K}}
\def\rank{\mathop{\rm rank}}
\def\half{{1/2}}
\def\Y{\mathbb{Y}}
\def\M{\mathbb{M}}
\def\F{\mathbb{F}}
\def\pinv{{-1}}
%\def\ones{\mathds{1}}
%\def\ones{1}
\def\Res{Z}
\def\Proj{P}
\def\cN{{\mathcal N}}
\def\cT{{\mathcal H}}
\def\coloneqq{:=}
\def\mathbf#1{\mbox{\boldmath $#1$}} 
\def\bar#1{\overline{#1}}

>>>>

% ENVIRONMENTS
% \numberwithin{equation}{section}
% \theoremstyle{plain}
% \newtheorem{theorem}{Theorem}[section]
% \newtheorem{corollary}{Corollary}[section]
% \newtheorem{proposition}{Proposition}[section]
% \newtheorem{lemma}{Lemma}[section]
% \newtheoremstyle{remark}{\topsep}{\topsep}%
%      {\normalfont}% Body font
%      {}           % Indent amount (empty = no indent, \parindent = para indent)
%      {\bfseries}  % Thm head font
%      {.}          % Punctuation after thm head
%      {.5em}       % Space after thm head (\newline = linebreak)
%      {\thmname{#1}\thmnumber{ #2}\thmnote{ #3}}% Thm head spec
% \theoremstyle{remark}
% \newtheorem{remark}{Remark}[section]
% \newtheorem{example}{Example}[section]
% \newtheorem{assumption}{Assumption}[section]
% \newtheorem{definition}{Definition}[section]

\input{macros}

@@

% John's macros
% \def\X{\mathcal{X}}
% \def\comma{\unskip,~}
% \def\truep{p^*}
% \def\div{\|\,}
% \long\def\comment#1{}
% \def\reals{{\mathbb R}}
% \def\P{{\mathbb P}}
% \def\E{{\mathbb E}}
% \def\Cov{\mathop{\text{Cov}}}
% \def\supp{\mathop{\text{supp}\kern.2ex}}
% \def\argmin{\mathop{\text{\rm arg\,min}}}
% \def\arginf{\mathop{\text{\rm arg\,inf}}}
% \def\argmax{\mathop{\text{\rm arg\,max}}}
% \let\tilde\widetilde
% \def\csd{${}^*$}
% \def\mld{${}^\dag$}
% \def\dos{${}^\ddag$}
% \def\W{\widetilde Y}
% \def\Z{\widetilde X}
% \let\hat\widehat
% \let\tilde\widetilde
% \def\given{{\,|\,}}
% \def\ds{\displaystyle}
% \def\bs{\backslash}
% \def\1{{(1)}}
% \def\2{{(2)}}
% \def\pn{{(n)}}
% \def\ip{{(i)}}
% \def\Xbar{\overline{X}}
% \def\except{\backslash}
% \def\npn{\mathop{\textit{NPN\,}}}
% \def\i{{(i)}}
% \def\cE{{\mathcal{C}}}
% \def\cM{{\mathcal{M}}}
% \def\cF{{\mathcal{F}}}
% \def\cP{{\mathcal{P}}}
% \def\cG{{\mathcal{G}}}
% \def\tr{\mathop{\text{tr}}}
% \long\def\comment#1{}
% \def\ti#1{#1}
% \def\titi#1{\textit{#1}}
% \def\cram{{\sc cram}}
% \def\spam{{\small\sc SpAM}}
% \def\diag{\mathop{\rm diag}}
% \def\ones{\mathbf{1}}
% \def\threebars{\mbox{$|\kern-.25ex|\kern-.25ex|$}}
% \def\fatnorm#1{\threebars #1 \threebars}
% \def\rank{\mathop{\rm rank}}
% \def\S{\mathcal{S}}
% \def\H{\mathcal{H}}
% \def\K{{K}}
% \def\rank{\mathop{\rm rank}}
% \def\half{{1/2}}
% \def\Y{\mathbb{Y}}
% \def\M{\mathbb{M}}
% \def\F{\mathbb{F}}
% \def\pinv{{-1}}
% %\def\ones{\mathds{1}}
% %\def\ones{1}
% \def\Res{Z}
% \def\Proj{P}
% \def\cN{{\mathcal N}}
% \def\cT{{\mathcal H}}
% \def\coloneqq{:=}
% \def\mathbf#1{\mbox{\boldmath $#1$}} 
% \def\bar#1{\overline{#1}}

### 2014-07-31 tex/scam/sparsistency.tex
minor typo
results update

\section{Analysis of Variable Selection Consistency}

@@

This result holds regardless of whether we impose the Lipschitz conditions in optimization~\ref{opt:alternate_opt} and~\ref{opt:alternate_opt_concave} or not.
The full proof of Theorem~\ref{thm:deterministic} is in Section~\ref{sec:deterministic_proof} of the Appendix.\\

@@

Each of our theorems will use a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent. 
\item[A2:] $f_0$ is convex and twice-differentiable. 
\item[A3:] $|\partial_{x_k} f_0 | \leq L$ for all $k$
\item[A4:] $w$ is mean-zero subgaussian, independent of $X$, with subgaussian scale $\sigma$, i.e. for all $t \in \R$, $\E e^{t \epsilon} \leq e^{\sigma^2 t^2 / 2}$.
\end{packed_enum}
By assumption A1, $f^*_k$ is must be zero for $k\notin \{1,...,s\}$.

@@

Suppose $\lambda_n \geq c s Lb \sigma  \sqrt{ \frac{1}{n} \log^2 np}$, then with probability at least $ 1 - \frac{C}{n}$, for all $k \in S^c$, and for all $i'=1,...,n$:
\[
\lambda_n > \Big| \frac{1}{2n}\hat{r}^\tran \mathbf{1}_{(i':n)} \Big|
\]

@@

\begin{theorem} (Controlling false negatives) \\
\label{thm:false_negative}
Suppose assumptions A1-A4 hold. Let $\hat{f}$ be any AC solution to the restricted regression with $L$-Lipschitz constraint and let $\hat{g}_k$'s be any DC solution to the restricted regression with $L$-Lipschitz constraint.\\

Suppose $\lambda \leq c sLb  \sqrt{\frac{1}{n} \log^2 np}$ and suppose $n$ is large enough such that $(Lb)^3 \sigma \sqrt{\frac{s^5}{n^{4/5}} \log^2 np} \leq 1$.\\

Suppose 
$\frac{\alpha_f}{\sigma} \geq c (Lb)^3 \sqrt{\frac{s^5}{n^{4/5}} \log^2 np}$ and $\frac{\alpha_g^2}{\sigma} \geq c (Lb)^3 \sqrt{\frac{s^5}{n^{4/5}} \log^2 2np}$.\\

@@

This is a finite sample version of
Theorem~\ref{thm:convex_faithful}. We need stronger assumptions in
Theorem~\ref{thm:false_negative} to use our additive faithfulness
result, Theorem~\ref{thm:convex_faithful}. We also include an extra
Lipschitz constraint so that we can use existing covering number
results \cite{Bronshtein:76}. Recent work
\cite{Guntu:13} shows that the Lipschitz constraint
is not required with more advanced empirical process theory
techniques; we leave the incorporation of this development as future work. 
We give the full proof of Theorem~\ref{thm:false_negative}
in Section~\ref{sec:false_negative_proof} of the Appendix.

Combining Theorem~\ref{thm:false_positive} and
~\ref{thm:false_negative} 
we have the following result.
\begin{corollary}
Suppose assumptions A1-A4 hold. Let $\hat{f}$ be any AC solution and $\hat{g}_k$'s be any DC solution, both with $L$-Lipschitz constraints, both with 
$\lambda = \Theta \Big( sLb \sqrt{\frac{1}{n} \log^2 np} \Big)$. \\

Suppose 
$\frac{\alpha_f}{\sigma} \geq c (Lb)^3 \sqrt{\frac{s^5}{n^{4/5}} \log^2 np}$ and $\frac{\alpha_g^2}{\sigma} \geq c (Lb)^3 \sqrt{\frac{s^5}{n^{4/5}} \log^2 2np}$.\\

@@

\begin{remark}
\citet{dalalyan:12} have shown that under tradtional smoothness constraints, variable selection is achievable only if $n > O(e^s)$. It is interesting to observe that because of additive faithfulness, the convexity assumption enables a much better scaling of $n = O(\textrm{poly}(s))$, demonstrating that geometric constraints can be quite different from the previously studied smoothness conditions.
\end{remark}

>>>>

\section{Analysis of Variable Selection Consistency}
\label{sec:finitesample}

@@

This result holds regardless of whether we impose the boundedness conditions in optimization~\ref{opt:alternate_opt} and~\ref{opt:alternate_opt_concave} or not.
The full proof of Theorem~\ref{thm:deterministic} is in Section~\ref{sec:deterministic_proof} of the Appendix.\\

@@

Each of our theorems will use a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent. 
\item[A2:] $f_0$ is convex and twice-differentiable. 
\item[A3:] $\|f_0\|_\infty \leq sB$ and $\| f^*_k \|_\infty \leq B$ for all $k$.
\item[A4:] $w$ is mean-zero subgaussian, independent of $X$, with subgaussian scale $\sigma$, i.e. for all $t \in \R$, $\E e^{t \epsilon} \leq e^{\sigma^2 t^2 / 2}$.
\end{packed_enum}
By assumption A1, $f^*_k$ is must be zero for $k\notin \{1,...,s\}$.

@@

Suppose $\lambda_n \geq c s B \sigma  \sqrt{ \frac{1}{n} \log^2 np}$, then with probability at least $ 1 - \frac{C}{n}$, for all $k \in S^c$, and for all $i'=1,...,n$:
\[
\lambda_n > \Big| \frac{1}{2n}\hat{r}^\tran \mathbf{1}_{(i':n)} \Big|
\]

@@

\begin{theorem} (Controlling false negatives) \\
\label{thm:false_negative}
Suppose assumptions A1-A4 hold. Let $\hat{f}$ be any AC solution to the restricted regression with $B$-boundedness and let $\hat{g}_k$'s be any DC solution to the restricted regression with $B$-boundedness constraint.\\

Suppose $\lambda \leq c sB  \sqrt{\frac{1}{n} \log^2 np}$ and $n$ is large enough such that$\frac{n^{4/5}}{\log^2 np} \geq B^3 \sigma s^{2.5}$.

Suppose 
$\frac{\alpha_f}{\sigma} \geq c B^3 \sqrt{\frac{s^5}{n^{4/5}} \log^2 np}$ and $\frac{\alpha_g^2}{\sigma} \geq c B^3 \sqrt{\frac{s^5}{n^{4/5}} \log^2 2np}$.\\

@@

This is a finite sample version of
Theorem~\ref{thm:convex_faithful}. We need stronger assumptions in
Theorem~\ref{thm:false_negative} to use our additive faithfulness
result, Theorem~\ref{thm:convex_faithful}. We also include an extra
boundedness constraint so that we can use recent bracketing number
results \cite{kim2014global}.

Combining Theorem~\ref{thm:false_positive} and
~\ref{thm:false_negative} 
we have the following result.
\begin{corollary}
Suppose assumptions A1-A4 hold. Let $\hat{f}$ be any AC solution and $\hat{g}_k$'s be any DC solution, both with $B$-boundedness constraints, both with 
$\lambda = \Theta \Big( sB \sqrt{\frac{1}{n} \log^2 np} \Big)$. \\

Suppose 
$\frac{\alpha_f}{\sigma} \geq c B^3 \sqrt{\frac{s^5}{n^{4/5}} \log^2 np}$ and $\frac{\alpha_g^2}{\sigma} \geq c B^3 \sqrt{\frac{s^5}{n^{4/5}} \log^2 2np}$.\\

@@

\begin{remark}
\citet{dalalyan:12} have shown that under tradtional smoothness constraints, even with a product distribution, variable selection is achievable only if $n > O(e^s)$. It is interesting to observe that because of additive faithfulness, the convexity assumption enables a much better scaling of $n = O(\textrm{poly}(s))$, demonstrating that geometric constraints can be quite different from the previously studied smoothness conditions.
\end{remark}

### 2014-07-31 tex/scam/summary.tex
results update

Each of our theorems uses a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent. 
\item[A2:] $f_0$ is convex and twice-differentiable. 
\item[A3:] $|\partial_{x_k} f_0 | \leq L$ for all $k$
\item[A4:] The noise is mean-zero sub-Gaussian, independent of $X$.
\end{packed_enum}

@@

We make strong assumptions on the covariates in A1 in order to make
very weak assumptions on the true regression function $f_0$ in
A2; in particular, we do not assume that $f_0$ is additive. 
Relaxing this condition is an important direction for future work.
We also include an extra
Lipschitz constraint so that we can use existing covering number
results \citep{Bronshtein:76}. Recent work
\cite{Guntu:13} shows that the Lipschitz constraint
is not required with more advanced empirical process theory
techniques; we leave the incorporation of this development as future work.

Our main result is the following.
\begin{stheorem}
Suppose assumptions A1-A4 hold. Let $\{\hat{f}_i\}$ be any AC solution and
let $\{\hat{g}_k\}$ be any DC solution, both estimated with 
regularization parameter $\lambda$ scaling as
$\lambda = \Theta \Big( sLb \sqrt{\frac{1}{n} \log^2 np} \Big)$. 
Suppose in addition that
\begin{gather}
\frac{\alpha_f}{\sigma} \geq c (Lb)^3 \sqrt{\frac{s^5}{n^{4/5}} \log^2
  np}\\
\frac{\alpha_g^2}{\sigma} \geq c (Lb)^3 \sqrt{\frac{s^5}{n^{4/5}}
  \log^2 2np}.
\end{gather}
Then, for sufficiently large $n$, with probability at least $1-\frac{1}{n}$:
\begin{align*}
\hat{f}_k \neq 0 \trm{ or } \hat{g}_k \neq 0 &\trm{ for all } k \in S\\
\hat{f}_k = 0 \trm{ and } \hat{g}_k = 0 & \trm{ for all } k \notin S.
\end{align*}

@@

We remark that \citet{dalalyan:12} show that under traditional smoothness
constraints, variable selection is achievable only if $n > O(e^s)$. 
Here we demonstrate that convexity yields the scaling $n =
O(\textrm{poly}(s))$.

>>>>

Each of our theorems uses a subset of the following assumptions:
\begin{packed_enum}
\item[A1:] $X_S, X_{S^c}$ are independent. 
\item[A2:] $f_0$ is convex and twice-differentiable. 
\item[A3:] $\|f_0\|_\infty \leq sB$ and $\|f^*_k \| \leq B$ for all $k$.
\item[A4:] The noise is mean-zero sub-Gaussian, independent of $X$.
\end{packed_enum}
In Assumption A3, $f^*=\sum_k f^*_k$ denotes the population optimal additive projection of $f_0$.

@@

We make strong assumptions on the covariates in A1 in order to make
very weak assumptions on the true regression function $f_0$ in
A2; in particular, we do not assume that $f_0$ is additive. 
Relaxing this condition is an important direction for future work.
We also include an extra
boundedness constraint so that we can use new bracketing number
results \citep{kim2014global}.

Our main result is the following.
\begin{stheorem}
Suppose assumptions A1-A4 hold. Let $\{\hat{f}_i\}$ be any AC solution and
let $\{\hat{g}_k\}$ be any DC solution, both estimated with 
regularization parameter $\lambda$ scaling as
$\lambda = \Theta \Big( sB \sqrt{\frac{1}{n} \log^2 np} \Big)$. 
Suppose in addition that
\begin{gather}
\frac{\alpha_f}{\sigma} \geq c B^3 \sqrt{\frac{s^5}{n^{4/5}} \log^2
  np}\\
\frac{\alpha_g^2}{\sigma} \geq c B^3 \sqrt{\frac{s^5}{n^{4/5}}
  \log^2 2np}.
\end{gather}
Then, for sufficiently large $n$, with probability at least $1-\frac{1}{n}$:
\begin{align*}
\hat{f}_k \neq 0 \trm{ or } \hat{g}_k \neq 0 &\trm{ for all } k \in S\\
\hat{f}_k = 0 \trm{ and } \hat{g}_k = 0 & \trm{ for all } k \notin S.
\end{align*}

@@

We remark that \citet{dalalyan:12} show that, even with the product distribution,
 under traditional smoothness
constraints, variable selection is achievable only if $n > O(e^s)$. 
Here we demonstrate that convexity yields the scaling $n =
O(\textrm{poly}(s))$.
